---
title: "Homework02"
author: "Avery Shepherd"
date: "3/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(cowplot)
```

## Challenge 01
```{r}
f <- "https://raw.githubusercontent.com/difiore/ADA-datasets/master/IMDB-movies.csv"
d <- read_csv(f,  col_names = T)
```

Use a one-line statement to filter the dataset to include just movies from 1920 to 1979 and movies that are between 1 and 3 hours long (runtimeMinutes < 240), and add a new column that codes the startYear into a new variable, decade (“20s,” “30s,” …“70s”). If you do this correctly, there should be 5651 movies remaining in the dataset.
```{r}
new_d <- d %>% filter(startYear >= 1920 & startYear <= 1979) %>% 
  filter(runtimeMinutes <= 180 & runtimeMinutes >= 60) %>% 
  mutate(decade = case_when(startYear >= 1920 & startYear < 1930 ~ "20s",
                          startYear >= 1930 & startYear < 1940 ~ "30s",
                          startYear >= 1940 & startYear < 1950 ~ "40s",
                          startYear >= 1950 & startYear < 1960 ~ "50s",
                          startYear >= 1960 & startYear < 1970 ~ "60s",
                          startYear >= 1970 & startYear < 1980 ~ "70s"))
new_d
```


```{r}
new_d %>% ggplot(aes(x = runtimeMinutes)) +
  geom_histogram() + 
  facet_wrap(~ decade)
```

Use a one-line statement to calculate the population mean and population standard deviation in runtimeMinutes for each decade and save the results in a new dataframe, results.
```{r}
pop_sd <- function(x) {
  sqrt((length(x)-1)/length(x)) * sd(x)
}
results <- new_d %>% group_by(decade) %>% summarise(runtime_mean = mean(runtimeMinutes), std_dev = pop_sd(runtimeMinutes), name = "pop")
results
```

Draw a single sample of 100 movies, without replacement, from each decade, calculate this single sample mean and sample standard deviation in runtimeMinutes, and estimate the SE around the population mean runtimeMinutes for each decade based on the standard deviation and sample size from these samples.
```{r}
set.seed(100)
sample_results <- new_d %>% group_by(decade) %>% sample_n(100) %>% summarise(runtime_mean = mean(runtimeMinutes), std_dev = sd(runtimeMinutes), name = "sample")
```

Compare these estimates to the actual population mean runtimeMinutes for each decade and to the calculated SE in the population mean for samples of size 100 based on the population standard deviation for each decade.
```{r}
p1 <- rbind(sample_results, results) %>% 
  ggplot(aes(x = name, y = runtime_mean)) + geom_point() + facet_wrap(~decade)
p2 <- rbind(sample_results, results) %>% 
  ggplot(aes(x = name, y = std_dev)) + geom_point() + facet_wrap(~decade)
plot_grid(p1, p2)
```

Generate a sampling distribution of mean runtimeMinutes for each decade by [a] drawing 1000 samples of 100 movies from each decade and, for each sample, [b] calculating the mean runtimeMinutes and the standard deviation in runtimeMinutes
```{r}
s <- data.frame()
for (i in 1:1000) {
  smple <- new_d %>% group_by(decade) %>% sample_n(100) %>% summarise(runtime_mean = mean(runtimeMinutes), std_dev = sd(runtimeMinutes), sample_num = i)
  s <- rbind(s, smple)
}
```

Then, calculate the MEAN and the STANDARD DEVIATION of the sampling distribution of sample means for each decade (the former should be a very good estimate of the population mean, while the latter is another estimate of the standard error in the population mean for a particular sample size)
```{r}
(distribution_results <- s %>% group_by(decade) %>% summarise(runtime_mean = mean(runtime_mean), std_dev = sd(std_dev), name = "distribution"))
```

plot a histogram of the sampling distribution. What shape does it have?
*normally distributed*
```{r}
ggplot(s, aes(x = runtime_mean)) + geom_histogram() + facet_wrap(~ decade)
```

Finally, compare the standard error in runtimeMinutes for samples of size 100 from each decade [1] as estimated from your first sample of 100 movies, [2] as calculated from the known population standard deviations for each decade, and [3] as estimated from the sampling distribution of sample means.
```{r}
standard_errors <- rbind(sample_results, results, distribution_results)
ggplot(data = standard_errors, aes(x = name, y = std_dev)) + 
  geom_point() + facet_wrap(~ decade)
```

## Challenge 02
Every morning at the same time, a bee bioloigist goes and sits in a field in the morning and watches for forager bees returning to a hive, counting the number that arrive in a one hour window from 7am to 8am. Based on previous knowledge, she believes that the mean number of foragers that will return in that time is 12, roughly one every 6 minutes. Let X represent the appropriate Poisson random variable of the number of arriving bees each monitoring session.

What is the probability that she will see 9 or fewer bees arrive during any given session?
```{r}
l = 6
x = c(0:9)
sum(dpois(x, l))
```

What is the probability that she will see no bees arrive in a session?
```{r}
x = 0
dpois(x, l)
```

What is the probability that she will see exactly 5 bees arrive in a session?
```{r}
x = 5
dpois(x, l)
```

What is the probability that she will see more than 18 bees arrive in a session?
```{r}
x = c(0:18)
1 - sum(dpois(x, l))
```

Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 24.
```{r}
x <- c(0:24)
prob <- dpois(x, l)
data.frame(cbind(x, prob)) %>% ggplot(aes(x, prob)) + geom_point()
```


```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2021-datasets/master/zombies.csv"
d <- read_csv(f, col_names = T)
```

Calculate the population mean and standard deviation for each quantitative random variable (height, weight, age, number of zombies killed, and years of education). Use {ggplot} and make boxplots of each of these variables by gender
```{r}
means <- d %>% summarise(m_height = mean(height), 
                m_weight = mean(weight), 
                m_age = mean(age), 
                m_kills = mean(zombies_killed), 
                m_edu = mean(years_of_education))
pop_sds <- d %>% summarise(sd_height = pop_sd(height), 
                sd_weight = pop_sd(weight),
                sd_age = pop_sd(age),
                sd_kills = pop_sd(zombies_killed),
                sd_edu = pop_sd(years_of_education))

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```